---
title: "Suchibrata"
author: "Suchibrata Patra - 403"
date: "2025-02-10"
output:
  word_document: default
  pdf_document: default
---

### part (a)
```{r}
rm(list=ls())
library(ggplot2)
library(tidyverse)
library(plotly)
data = read.csv("data.csv") ; data_scaled = scale(data[-1]) # Used this to 
cov(data[-1]) # variability in the expenditure is varying w.r.t Food items
cor(data[-1])
```

###Comments from Covariance Matrix
Matrix is Symmetric.
Meat seems to have highest Variation in Consumption Value.

### Comments from Correlatino Matrix
Matrix is symmetric.
Meat and poultry items seems to have High Correlation values.
Wine have Negative Correlation with Some of the food items.



```{r}

cov_matrix = cov(data_scaled)
eigen_decomp = eigen(cov_matrix)
eigenvalues = eigen_decomp$values
eigenvectors = eigen_decomp$vectors

# Selecting top 2 components
selected_components = eigenvectors[, 1:2]

# Transforming the data
pca_data = as.data.frame(data_scaled %*% selected_components)
colnames(pca_data) = c("x1", "x2")

# Add first column of original DataSet as labels
pca_data$label = data[, 1]

p = ggplot(pca_data, aes(x = x1, y = x2)) +
  geom_point(color = "coral", alpha = 2) +
  geom_text(aes(label = label),size = 3,nudge_y = 0.2) +
  labs(title = "Reduced Dimension",
       x = "Component 1",
       y = "Component 2") +
  theme_minimal()

final_plot = ggplotly(p) ; final_plot
#I have added 'ggplotly()' So that tooltip is shown over there when the points are hovered.

```


# Comments 

The x-axis and y-axis represent the first(x_1) and second principal components(x_2), respectively.
These components are linear combinations of the original features that maximize variance.

A specific data point labeled "CA5" is plotted near the origin on Component 1 (x-axis) and Component 2 (y-axis).
The spread of points indicate variance explained by each principal component.
Since 'CA5' is close to the origin, it suggests that it does not strongly contribute to variance along these components.



